---
output: github_document
bibliography: vignettes/biblio_phd_zot.bib
csl: vignettes/emerald-harvard.csl
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Fluxible

<!-- badges: start -->
[![R-CMD-check](https://github.com/Plant-Functional-Trait-Course/fluxible/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Plant-Functional-Trait-Course/fluxible/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/fluxible)](https://CRAN.R-project.org/package=fluxible)
[![LifeCycle](https://img.shields.io/badge/lifecycle-stable-green)](https://lifecycle.r-lib.org/articles/stages.html#stable)
<!-- badges: end -->

The Fluxible R package is made to transform any dataset of gas concentration over time into a gas flux dataset.
It was originally made to be used with a closed loop chamber system connected to a gas analyzer.

The goal of fluxible is to provide a workflow that removes individual evaluation of each flux, reduces risk of bias, and makes it reproducible.
Users set specific data quality standards and selection parameters as function arguments that are applied to the entire dataset.
Fluxible offers different methods to estimate fluxes: linear, quadratic, exponential [@zhaoCalculationDaytimeCO22018], and the original HM model [@hutchinsonImprovedSoilCover1981; @pedersenComprehensiveApproachSoilatmosphere2010a].
The kappamax method [@huppiRestrictingNonlinearityParameter2018] is also included, at the quality control step.
The package runs the calculations automatically, without prompting the user to take decisions mid-way, and provides quality flags and plots at the end of the process for a visual check.

This makes it easy to use with large flux datasets and to integrate into a reproducible and automated data processing pipeline such as the [targets R package [@targetsRpackage2021]](https://books.ropensci.org/targets/).
Using the Fluxible R package makes the workflow reproducible, increases compatibility across studies, and is more time efficient.

For a visual overview of the package, see [the poster](https://raw.githubusercontent.com/Plant-Functional-Trait-Course/fluxible/refs/heads/main/dissemination/poster_fluxible.pdf).

## Installation

Fluxible can be installed from CRAN.

``` r
install.packages("fluxible")
```


You can install the development version of fluxible from the [GitHub repo](https://github.com/Plant-Functional-Trait-Course/fluxible) with:

``` r
# install.packages("devtools")
devtools::install_github("plant-functional-trait-course/fluxible")
```

## Short example
```{r short-example, fig.width = 8, fig.height = 9, fig.cap="Output of flux_plot, showing fluxes plotted individually with diagnostics and quality flags."}
library(fluxible)

conc_df <- flux_match(
  co2_df_short,
  record_short,
  datetime,
  start,
  measurement_length = 220
)

slopes_df <- flux_fitting(
  conc_df,
  conc,
  datetime,
  fit_type = "exp_zhao18",
  end_cut = 60
)

slopes_flag_df <- flux_quality(
  slopes_df,
  conc
)

flux_plot(
  slopes_flag_df,
  conc,
  datetime,
  f_ylim_lower = 390,
  f_ylim_upper = 650,
  facet_wrap_args = list(
    ncol = 3,
    nrow = 2,
    scales = "free"
  )
)

fluxes_df <- flux_calc(
  slopes_flag_df,
  f_slope_corr,
  datetime,
  temp_air,
  conc_unit = "ppm",
  flux_unit = "mmol",
  cols_keep = c("turfID", "type"),
  cols_ave = c("temp_soil", "PAR"),
  setup_volume = 24.575,
  atm_pressure = 1,
  plot_area = 0.0625
)

fluxes_gpp <- flux_gpp(
  fluxes_df,
  type,
  datetime,
  id_cols = "turfID",
  cols_keep = c("temp_soil_ave")
)

fluxes_gpp
```


## Further developments

### The licoread R package

The [licoread R package](https://joXXXX.github.io/licoread/index.html), developped in collaboration with [LI-COR](https://www.licor.com/), provides an easy way to import raw files from Li-COR gas analyzers as R objects that can be used directly with the Fluxible R package.

### Segmentation tool
We are working on a tool to automatically select the window of the measurement on which to fit a model.
This selection will be based on environmental variable, such as photosynthetically active radiation (PAR), measured simultaneously.

### More fits
As we want fluxible to fit the use of as many projects as possible, more fitting expressions will be included in the flux_fitting function.
Feel welcome to get in touch if you wish to include yours in fluxible.

## Contact
[XXXX XXXX](https://joXXXX.github.io/CV_jXXXX/), University of XXXX, XXXX

[XXXX.XXXX@pm.me](mailto:XXXX.XXXX@pm.me)

[GitHub page](https://github.com/joXXXX)

## Dissemination

If you are running a course and want to talk about Fluxible, feel free to use this [two-slides presentation](https://raw.githubusercontent.com/Plant-Functional-Trait-Course/fluxible/refs/heads/main/dissemination/fluxible_jXXXX_short.pdf).
Of course, you can always reach out if you wish to have more material.

XXXX, J., XXXX, J., XXXX, H. R., XXXX, B., XXXX, A. H., Mustri, M., Niittynen, P., XXXX, P. E., XXXX, J. P., XXXX, V., and XXXX, R. J.: Fluxible: an R package to calculate ecosystem gas fluxes from closed loop chamber systems in a reproducible and automated workflow ([slides](https://raw.githubusercontent.com/Plant-Functional-Trait-Course/fluxible/refs/heads/main/dissemination/fluxible_jXXXX_EGU25.pdf)), EGU General Assembly 2025, Vienna, Austria, 27 Apr–2 May 2025, EGU25-12409, https://doi.org/10.5194/egusphere-egu25-12409, 2025. 

XXXX, J., XXXX, R., XXXX, V., and XXXX, A. H.: Fluxible: an R package to calculate ecosystem gas fluxes in a reproducible and automated workflow. ([poster](https://raw.githubusercontent.com/Plant-Functional-Trait-Course/fluxible/refs/heads/main/dissemination/poster_fluxible.pdf)), EGU General Assembly 2024, Vienna, Austria, 14–19 Apr 2024, EGU24-956, https://doi.org/10.5194/egusphere-egu24-956, 2024. 

<!-- [Direct link to the poster](https://github.com/Plant-Functional-Trait-Course/fluxible/blob/main/dissemination/poster_EGU24_jXXXX.pdf) -->

<!-- [Link to the abstract](https://meetingorganizer.copernicus.org/EGU24/EGU24-956.html) -->


## Acknowledgements

Fluxible builds on the earlier effort from the Plant Functional Traits Course Community [co2fluxtent](https://github.com/XXXXESantos/co2fluxtent) [@co2fluxtent].

#### References
