---
output: github_document
bibliography: vignettes/biblio_phd_zot.bib
csl: vignettes/emerald-harvard.csl
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Fluxible

**Note:** fluxible is still in an early phase, it is fully functional but can be improved in many ways. Please get in touch for any problems or suggestions.

<!-- badges: start -->
<!-- badges: end -->
The Fluxible R package is made to transform any dataset of gas concentration over time into a gas flux dataset.
It was originally made to be used with a closed loop chamber system connected to a gas analyzer.
A function for use with a tent (taking leakage into account) is in development (see further developments).

The goal of fluxible is to provide a workflow that removes individual evaluation of each flux, reduces risk of bias, and makes it reproducible.
Users set specific data quality standards and selection parameters as function arguments that are applied to the entire dataset.
The package runs the calculations automatically, without prompting the user to take decisions mid-way, and provides quality flags and plots at the end of the process for a visual check.
This makes it easy to use with large flux datasets and to integrate into a reproducible workflow.
Using the Fluxible R package makes the workflow reproducible, increases compatibility across studies, and is more time efficient.

For a quick overview of how the package works, you can look at [the poster](https://github.com/Plant-Functional-Trait-Course/fluxible/blob/main/dissemination/poster_EGU24_jgaudard.pdf) that was presented at the European Geosciences Union Annual Meeting 2024.

## Installation

You can install the development version of fluxible from the [GitHub repo](https://github.com/Plant-Functional-Trait-Course/fluxible) with:

``` r
# install.packages("devtools")
devtools::install_github("plant-functional-trait-course/fluxible")
```

## Example
For this example we will use the data that were recorded during the Plant Functional Traits Course 6 (PFTC6) in Norway in 2022 at the site called Liahovden. 
<!-- (CITE when data paper out). -->
The CO~2~ concentration data as well as air and soil temperature and photosynthetically active radiations (PAR) were recorded in a dataframe named `co2_liahovden`.
The metadata for each measurements are in a dataframe called `record_liahovden`.
This dataframe contains the starting time of each measurements, the type of measurement and the unique ID for each turf.
The type of measurement describes if it was net ecosystem exchange (NEE), measured with a transparent chamber, or ecosystem respiration (ER), measured with a dark chamber.

We use the `flux_match` function to slice the data from `co2_liahovden` into each measurement and discard what was recorded in between.


```{r match}
library(fluxible)

str(record_liahovden)
str(co2_liahovden)

conc_liahovden <- flux_match(
  raw_conc = co2_liahovden,
  field_record = record_liahovden,
  startcrop = 0,
  measurement_length = 220,
  ratio_threshold = 0.5,
  time_diff = 0,
  datetime_col = "datetime",
  conc_col = "conc",
  start_col = "start"
)
str(conc_liahovden)
```

Before calculating fluxes we need to fit a model to each measurement and estimate a slope of the concentration changing rate.
We use the `flux_fitting` function with the model provided by @zhaoCalculationDaytimeCO22018.
The function `flux_fitting` also provides a quadratic and a linear fit.


```{r fitting_exp_qua_lin}
slopes_exp_liahovden <- flux_fitting(
  conc_df = conc_liahovden,
  start_cut = 0,
  end_cut = 0,
  start_col = "f_start",
  end_col = "f_end",
  datetime_col = "f_datetime",
  conc_col = "f_conc",
  fluxid_col = "f_fluxID",
  t_window = 20,
  cz_window = 15,
  b_window = 10,
  a_window = 10,
  roll_width = 15,
  t_zero = 0,
  fit_type = "exponential"
)
str(slopes_exp_liahovden)

slopes_qua_liahovden <- flux_fitting(
  conc_df = conc_liahovden,
  start_cut = 0,
  end_cut = 0,
  start_col = "f_start",
  end_col = "f_end",
  datetime_col = "f_datetime",
  conc_col = "f_conc",
  fluxid_col = "f_fluxID",
  t_window = 20,
  cz_window = 15,
  b_window = 10,
  a_window = 10,
  roll_width = 15,
  t_zero = 5,
  fit_type = "quadratic"
)
str(slopes_qua_liahovden)

slopes_lin_liahovden <- flux_fitting(
  conc_df = conc_liahovden,
  start_cut = 0,
  end_cut = 0,
  start_col = "f_start",
  end_col = "f_end",
  datetime_col = "f_datetime",
  conc_col = "f_conc",
  fluxid_col = "f_fluxID",
  t_window = 20,
  cz_window = 15,
  b_window = 10,
  a_window = 10,
  roll_width = 15,
  t_zero = 5,
  fit_type = "linear"
)
str(slopes_lin_liahovden)
```

The function `flux_quality` is used to provide diagnostics about the quality of the fit, eventually advising to discard some measurements or replace them by zero.

```{r quality_exp_qua_lin}
slopes_exp_liahovden_flag <- flux_quality(
  slopes_df = slopes_exp_liahovden,
  # fit_type is automatically provided as an attribute because
  # slopes_exp_liahovden was produced with flux_fitting
  ambient_conc = 421,
  error = 100,
  fluxid_col = "f_fluxID",
  slope_col = "f_slope",
  weird_fluxes_id = c(),
  force_ok_id = c(),
  ratio_threshold = 0,
  conc_col = "f_conc",
  b_col = "f_b",
  time_col = "f_time",
  fit_col = "f_fit",
  cut_col = "f_cut",
  rmse_threshold = 25,
  cor_threshold = 0.5,
  b_threshold = 1,
  cut_arg = "cut"
)
str(slopes_exp_liahovden_flag)

slopes_qua_liahovden_flag <- flux_quality(
  slopes_df = slopes_qua_liahovden,
  # fit_type is automatically provided as an attribute because
  # slopes_exp_liahovden was produced with flux_fitting
  ambient_conc = 421,
  error = 100,
  fluxid_col = "f_fluxID",
  slope_col = "f_slope",
  weird_fluxes_id = c(),
  force_ok_id = c(),
  ratio_threshold = 0,
  pvalue_col = "f_pvalue",
  rsquared_col = "f_rsquared",
  pvalue_threshold = 0.3,
  rsquared_threshold = 0.7,
  conc_col = "f_conc",
  time_col = "f_time",
  fit_col = "f_fit",
  cut_col = "f_cut",
  cut_arg = "cut"
)
str(slopes_qua_liahovden_flag)

slopes_lin_liahovden_flag <- flux_quality(
  slopes_df = slopes_lin_liahovden,
  # fit_type is automatically provided as an attribute because
  # slopes_exp_liahovden was produced with flux_fitting
  ambient_conc = 421,
  error = 100,
  fluxid_col = "f_fluxID",
  slope_col = "f_slope",
  weird_fluxes_id = c(),
  force_ok_id = c(),
  ratio_threshold = 0,
  pvalue_col = "f_pvalue",
  rsquared_col = "f_rsquared",
  pvalue_threshold = 0.3,
  rsquared_threshold = 0.7,
  conc_col = "f_conc",
  time_col = "f_time",
  fit_col = "f_fit",
  cut_col = "f_cut",
  cut_arg = "cut"
)
str(slopes_lin_liahovden_flag)
```

The function `flux_plot` provides plots for a visual assessment of the measurements, explicitly displaying the quality flags from `flux_quality` and the cuts from `flux_fitting`.

```{r plot_exp_qua_lin, fig.width = 8, fig.height = 9}
slopes_exp_liahovden_flag |>
  # we just show a sample of the plots to avoid slowing down the example
  dplyr::filter(f_fluxID %in% c(54, 95, 100, 101)) |>
  flux_plot(
    color_discard = "#D55E00",
    color_cut = "#D55E00",
    color_ok = "#009E73",
    color_zero = "#CC79A7",
    f_date_breaks = "1 min",
    f_minor_breaks = "10 sec",
    f_date_labels = "%e/%m \n %H:%M",
    f_ylim_upper = 600,
    f_ylim_lower = 300,
    f_plotname = "plot_quality",
    facet_wrap_args = list(
      ncol = 2,
      nrow = 2,
      scales = "free"
    ),
    y_text_position = 400,
    print_plot = "FALSE",
    output = "print_only",
    cut_arg = "cut",
    no_data_flag = "no_data"
  )

slopes_qua_liahovden_flag |>
  # we just show a sample of the plots to avoid slowing down the example
  dplyr::filter(f_fluxID %in% c(54, 95, 100, 101)) |>
  flux_plot(
    color_discard = "#D55E00",
    color_cut = "#D55E00",
    color_ok = "#009E73",
    color_zero = "#CC79A7",
    f_date_breaks = "1 min",
    f_minor_breaks = "10 sec",
    f_date_labels = "%e/%m \n %H:%M",
    f_ylim_upper = 600,
    f_ylim_lower = 300,
    f_plotname = "plot_quality",
    facet_wrap_args = list(
      ncol = 2,
      nrow = 2,
      scales = "free"
    ),
    y_text_position = 400,
    print_plot = "FALSE",
    output = "print_only",
    cut_arg = "cut",
    no_data_flag = "no_data"
  )

slopes_lin_liahovden_flag |>
  # we just show a sample of the plots to avoid slowing down the example
  dplyr::filter(f_fluxID %in% c(54, 95, 100, 101)) |>
  flux_plot(
    color_discard = "#D55E00",
    color_cut = "#D55E00",
    color_ok = "#009E73",
    color_zero = "#CC79A7",
    f_date_breaks = "1 min",
    f_minor_breaks = "10 sec",
    f_date_labels = "%e/%m \n %H:%M",
    f_ylim_upper = 600,
    f_ylim_lower = 300,
    f_plotname = "plot_quality",
    facet_wrap_args = list(
      ncol = 2,
      nrow = 2,
      scales = "free"
    ),
    y_text_position = 400,
    print_plot = "FALSE",
    output = "print_only",
    cut_arg = "cut",
    no_data_flag = "no_data"
  )
```

Based on the quality flags and the plots, the user can decide to run `flux_fitting` again with different arguments.
Here we will do it while cutting the last 60 seconds of the fluxes (cutting the last third).
We also detected fluxes that do not look correct.
Sometimes some measurements will pass the automated quality control but are obviously wrong for an experience user.
That is what the `weird_fluxesID` argument is for.
For the sake of reproducibility, this argument should be the last option and be accompanied with a justification.

```{r plot_exp_cut, fig.width = 8, fig.height = 9}
slopes_exp_liahovden_flag_60 <- conc_liahovden |>
  flux_fitting(fit_type = "exp", end_cut = 60) |>
  flux_quality(
    slope_col = "f_slope",
    weird_fluxes_id = c(
      51, # slope is much steeper than the flux because t zero was estimated
      # at the very start of the measurement
      101, # plot starts with a high peak: accumulation in the canopy?
      106 # peak at the beginning of the flux that is messing up the fit
    )
  )

slopes_exp_liahovden_flag_60 |>
  # we just show a sample of the plots to avoid slowing down the example
  dplyr::filter(f_fluxID %in% c(54, 95, 100, 101)) |>
  flux_plot(
    f_ylim_lower = 300,
    f_ylim_upper = 600,
    facet_wrap_args = list(
      ncol = 2,
      nrow = 2,
      scales = "free"
    ),
    y_text_position = 400,
    output = "print_only"
  )
```

We also apply a cut on the dataset that was fitted with a quadratic model.
At this point it is up to the user to decide which model works the best for the entire dataset.
The function flux_quality provides a count of the quality flags that can help to take a decision.

```{r plot_qua_cut, fig.width = 8, fig.height = 9}
slopes_qua_liahovden_flag_60 <- conc_liahovden |>
  flux_fitting(fit_type = "qua", end_cut = 60, t_zero = 5) |>
  flux_quality(
    slope_col = "f_slope"
  )

slopes_qua_liahovden_flag_60 |>
  # we just show a sample of the plots to avoid slowing down the example
  dplyr::filter(f_fluxID %in% c(54, 95, 100, 101)) |>
  flux_plot(
    f_ylim_lower = 300,
    f_ylim_upper = 600,
    facet_wrap_args = list(
      ncol = 2,
      nrow = 2,
      scales = "free"
    ),
    y_text_position = 400,
    output = "print_only"
  )
```

When using a linear fit it is common to take only a short section of the measurement close to the start.
Here we will cut 120 seconds at the end, effectively keeping only the first 90 seconds.

```{r plot_lin_cut, fig.width = 8, fig.height = 9}
slopes_lin_liahovden_flag_120 <- conc_liahovden |>
  flux_fitting(fit_type = "lin", end_cut = 120, t_zero = 5) |>
  flux_quality(
    slope_col = "f_slope"
  )

slopes_lin_liahovden_flag_120 |>
  # we just show a sample of the plots to avoid slowing down the example
  dplyr::filter(f_fluxID %in% c(54, 95, 100, 101)) |>
  flux_plot(
    f_ylim_lower = 300,
    f_ylim_upper = 600,
    facet_wrap_args = list(
      ncol = 2,
      nrow = 2,
      scales = "free"
    ),
    y_text_position = 400,
    output = "print_only"
  )
```

Once we are satisfied with the fit, we can calculate fluxes with `flux_calc`.
Here the volume is defined as a constant for all the measurements but it is also possible to provide a specific volume for each plot in case that is different.

```{r calc}
fluxes_exp_liahovden_60 <- slopes_exp_liahovden_flag_60 |>
  flux_calc(
    slope_col = "f_slope_corr", # we use the slopes provided by flux_quality
    datetime_col = "f_datetime",
    cut_col = "f_cut",
    keep_arg = "keep",
    chamber_volume = 24.5,
    tube_volume = 0.075,
    atm_pressure = 1,
    plot_area = 0.0625,
    cols_keep = c("f_start", "type"),
    cols_ave = c(),
    fluxid_col = "f_fluxID",
    temp_air_col = "temp_air",
    temp_air_unit = "celsius"
  )
str(fluxes_exp_liahovden_60)

fluxes_qua_liahovden_60 <- slopes_qua_liahovden_flag_60 |>
  flux_calc(
    slope_col = "f_slope_corr", # we use the slopes provided by flux_quality
    datetime_col = "f_datetime",
    cut_col = "f_cut",
    keep_arg = "keep",
    chamber_volume = 24.5,
    tube_volume = 0.075,
    atm_pressure = 1,
    plot_area = 0.0625,
    cols_keep = c("f_start", "type"),
    cols_ave = c(),
    fluxid_col = "f_fluxID",
    temp_air_col = "temp_air",
    temp_air_unit = "celsius"
  )
str(fluxes_qua_liahovden_60)

fluxes_lin_liahovden_120 <- slopes_lin_liahovden_flag_120 |>
  flux_calc(
    slope_col = "f_slope_corr", # we use the slopes provided by flux_quality
    datetime_col = "f_datetime",
    cut_col = "f_cut",
    keep_arg = "keep",
    chamber_volume = 24.5,
    tube_volume = 0.075,
    atm_pressure = 1,
    plot_area = 0.0625,
    cols_keep = c("f_start", "type"),
    cols_ave = c(),
    fluxid_col = "f_fluxID",
    temp_air_col = "temp_air",
    temp_air_unit = "celsius"
  )
str(fluxes_lin_liahovden_120)
```

The output is in mmol/m^2^/h and the calculation used is as follow:

<!-- <img src="https://render.githubusercontent.com/render/math?math=flux=slope\times \frac{P\times V}{R\times T\times A}"> -->

$$
 \text{flux}=\text{slope}\times \frac{P\times V}{R\times T\times A}
$$

where

flux: the flux of gas at the surface of the plot (mmol/m^2^/h)

slope: slope estimate (ppm*s^-1^)

P: pressure, assumed  (atm)

V: volume of the chamber and tubing (L)

R: gas constant (0.082057 L\*atm\*K^-1^\*mol^-1^)

T: chamber air temperature (K)

A: area of chamber frame base (m^2^)

The conversion from micromol/m^2^/s to mmol/m^2^/h is included in the function.

Fluxes were calculated in five steps from raw gas concentration data and the process is entirely reproducible.
Here is a plot of the results.

```{r 24h_fluxes, fig.width = 8, fig.height = 9, warning = FALSE, message = FALSE}
library(dplyr)
library(ggplot2)
bind_rows(
  fluxes_exp_liahovden_60,
  fluxes_qua_liahovden_60,
  fluxes_lin_liahovden_120
) |>
  ggplot(aes(x = f_start, y = flux, color = model)) +
  geom_point() +
  geom_smooth() +
  labs(
    title = "Net Ecosystem Exchange at Upper Site (Liahovden) during 24 hour",
    x = "Datetime",
    y = bquote(~ CO[2] ~ "flux [mmol/" * m^2 * "/h]"),
    color = "Model used in flux_fitting"
  ) +
  theme(legend.position = "bottom") +
  facet_grid(type ~ ., scales = "free")
```


## Further developments

### Flux tent
Currently fluxible is thought for setups using a flux chamber.
Flux chambers can be assumed to have no leak.
As flux tents tend to leak, a different fitting function is needed to account for potential leaks.

### More fits
As we want fluxible to fit the use of as many projects as possible, more fitting expressions will be included in the flux_fitting function.
Feel welcome to get in touch if you wish to include yours in fluxible.

## Contact
Joseph Gaudard, University of Bergen, Norway

joseph.gaudard@pm.me

[CV](https://jogaudard.github.io/CV_jgaudard/)

## Dissemination
Gaudard, J., Telford, R., Vandvik, V., and Halbritter, A. H.: Fluxible: an R package to calculate ecosystem gas fluxes in a reproducible and automated workflow., EGU General Assembly 2024, Vienna, Austria, 14–19 Apr 2024, EGU24-956, https://doi.org/10.5194/egusphere-egu24-956, 2024. 

[Direct link to the poster](https://github.com/Plant-Functional-Trait-Course/fluxible/blob/main/dissemination/poster_EGU24_jgaudard.pdf)

<!-- [Link to the abstract](https://meetingorganizer.copernicus.org/EGU24/EGU24-956.html) -->

#### References
