---
title: "Using fluxible with the Li7500"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Preparing the data for fluxible}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: biblio_phd_zot.bib
csl: emerald-harvard.csl
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(tidyverse.quiet = TRUE)
```

In this vignette, we will see how to read the data from the setup described in @halbritterPlantTraitVegetation2024 and compare the output calculated with fluxible with the published data.

The code used to import the data might be turned into a package later.

## Importing and reading the files

First, we import the data from OSF with `dataDownloader::get_file`.

```{r osf-import, include = TRUE}
library(dataDownloader)

get_file("gs8u6",
         "rawData/raw_c-flux",
         "rawC-flux.zip",
         "vignettes/ex_data/li7500")

unzip("vignettes/ex_data/li7500/rawC-flux.zip",
      exdir = "vignettes/ex_data/li7500/rawC-flux")
```

Then we import the data with `licoread::import7500`.
```{r reading}
library(licoread)
pftc5_data <- import7500(
  "vignettes/ex_data/li7500/rawC-flux/rawData/raw_tent-flux",
  plotinfo_names = c("site", "block", "date", "plot", "trial")
)
```


## Processing with fluxible
Since the `licoread::import7500` already provides `f_fluxid`, `f_start` and `f_end`, we can skip `fluxible::flux_match` and directly fit with `fluxible::flux_fitting`.

```{r fitting}
library(fluxible)
pftc5_fits <- flux_fitting(pftc5_data,
                           f_conc = `CO2 umol/mol`,
                           fit_type = "exp_zhao18")
```

Using `fluxible::flux_quality` to assess the quality of the dataset.
```{r flags}
pftc5_flags <- flux_quality(pftc5_fits,
                            f_conc = `CO2 umol/mol`)
```

```{r sites, include = FALSE}
pftc5_flags |>
  dplyr::select(site) |>
  dplyr::distinct()
```

Then plotting (in an external file) for a visual check.
Because of the number of data, we will do it per site to make it lighter.
```{r plotting, eval = FALSE}
pftc5_flags |>
  dplyr::filter(site == "ACJ") |>
  flux_plot(f_conc = `CO2 umol/mol`,
            print_plot = FALSE,
            output = "longpdf",
            f_plotname = "ACJ_plot")

pftc5_flags |>
  dplyr::filter(site == "PIL") |>
  flux_plot(f_conc = `CO2 umol/mol`,
            print_plot = FALSE,
            output = "longpdf",
            f_plotname = "PIL_plot")

pftc5_flags |>
  dplyr::filter(site == "QUE") |>
  flux_plot(f_conc = `CO2 umol/mol`,
            print_plot = FALSE,
            output = "longpdf",
            f_plotname = "QUE_plot")

pftc5_flags |>
  dplyr::filter(site == "TRE") |>
  flux_plot(f_conc = `CO2 umol/mol`,
            print_plot = FALSE,
            output = "longpdf",
            f_plotname = "TRE_plot")

pftc5_flags |>
  dplyr::filter(site == "WAY") |>
  flux_plot(f_conc = `CO2 umol/mol`,
            print_plot = FALSE,
            output = "longpdf",
            f_plotname = "WAY_plot")
```

Now let's calculate the fluxes with `fluxible::flux_calc`.
```{r calc}
pftc5_fluxes <- flux_calc(pftc5_flags,
                          slope_col = f_slope_corr,
                          temp_air_col = Temperature,
                          setup_volume = 2197,
                          atm_pressure = pressure_atm,
                          plot_area = 1.44,
                          conc_unit = "ppm",
                          flux_unit = "umol/m2/s")
```

## Comparison

### Importing and reading published data
