In this vignette, we will process raw data both with Fluxible and the gasfluxes package [@fussGasfluxesGreenhouseGas2024], then compare the output.
This is also a good opportunity to compare the workflow.

```{r}
#| label: data-process-stupeflux
#| echo: TRUE
library(tidyverse)
library(fluxible)

# we can use the stupeflux function, which is wrapper for fluxible
# we use the arguments that we already know from having processed
# those data in the detailed example.

fluxible_fluxes_hm <- stupeflux(
    raw_conc = co2_liahovden,
    field_record = record_liahovden,
    datetime_col = datetime,
    start_col = start,
    conc_col = conc,
    startcrop = 0,
    measurement_length = 180,
    fit_type = "exp_hm",
    end_cut = 10,
    force_discard = c(51, 101),
    force_zero = 46,
    temp_air_col = temp_air,
    conc_unit = "ppm",
    flux_unit = "mmol",
    chamber_volume = 24.5,
    tube_volume = 0.075,
    atm_pressure = 1,
    plot_area = 0.0625,
    cols_keep = "type"
 ) |>
 select(f_fluxid, f_flux, type) |>
 rename(
    fluxible_flux = "f_flux"
 )
```


Now let's dive into the gasfluxes package.
We will use the `kappa.max` method [@huppiRestrictingNonlinearityParameter2018] as this is the most recent one.

```{r}
#| label: data-process-gasfluxes
#| echo: TRUE
library(gasfluxes)
# library(lubridate)

# we use the same data from fluxible
# we use fluxible::flux_match because gasfluxes does not provide such a function

conc_lia <- fluxible::flux_match(
  raw_conc = co2_liahovden,
  field_record = record_liahovden,
  datetime_col = datetime,
  start_col = start,
  conc_col = conc,
  startcrop = 0,
  measurement_length = 170, # gasfluxes does not provide a cutting tool, so we just do it here
  ratio_threshold = 0.5,
  time_diff = 0
)

# gasfluxes requires volumetric concentration and not fractional

Rconst <- 8.20573660809596e-2 # gas constant in m³ * atm * K⁻¹ * mmol ⁻¹
atmP <- 1 # atmospheric pressure assumed 1 atm

conc_lia_vol <- conc_lia |>
    group_by(f_fluxid) |>
    fill(temp_air, .direction = "downup") |> # air temperature was measured every 10 seconds
    mutate(
        # temp_air = 
        temp_air = temp_air + 273.15, # we want air temperature in Kelvin
        C = conc * (atmP / (Rconst * temp_air)) # gas concentration in mol / m³
    )

# gasfluxes requires columns with chamber volume (V), plot area (A), and elapsed time (time)

conc_lia_gasfluxes <- conc_lia_vol |>
    rename(
        ID = "f_fluxid"
    ) |>
    mutate(
        V = 0.024575, # total of chamber and tubes, in m3
        A = 0.0625, # in m2
        time = difftime(datetime[seq_along(datetime)],
            datetime[1],
            units = "hours" # gasfluxes wants time in hours, which might be suboptimal since most modern gas analyzers measure concentration every second
      ),
      time = as.numeric(time)
    ) |>
    select(ID, V, A, time, C)

# now the data are ready for gasfluxes

# it prints all the flux ID with "sucessful" or "failed" and I do not like test
qgasfluxes <- quietly(gasfluxes)


gasfluxes_fits <- qgasfluxes(conc_lia_gasfluxes, method = c("linear","robust linear", "HMR"), plot = TRUE)

# str(gasfluxes_fits)
# the output is a list, I might have an opinion about this

# now let's try to select the flux estimates with the kappa.max method
gasfluxes_fluxes <- selectfluxes(
                        gasfluxes_fits$result, # that is not specified in the vignette
                        select = "kappa.max",
                        f.detect = 0.031, # I keep the suggested value for now
                        t.meas = 0.05 # I think this is the length of the measurement in hours
                    ) |>
                    select(ID, flux, method, flux.se) |>
                    rename(
                        gasfluxes_flux = "flux"
                    )

```

Now we can compare.

```{r}
#| label: gasfluxes-comparison
#| echo: TRUE
#| fig-width: 8
#| fig-height: 9
#| fig-cap: ""

library(ggplot2)

all_fluxes <- left_join(gasfluxes_fluxes, fluxible_fluxes, by = c("ID" = "f_fluxid"))

all_fluxes |>
    ggplot(aes(gasfluxes_flux, fluxible_flux, colour = method)) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0) +
    geom_smooth(
        method = "lm"
        ) +
    geom_errorbar(aes(xmin = gasfluxes_flux - flux.se, xmax = gasfluxes_flux + flux.se)) +
    facet_grid(~type)

# investigating that flux with a massive difference

# slice_max(all_fluxes, fluxible_flux)

# flags_lia |>
# filter(f_fluxid == 46) |>
# flux_plot(conc, datetime)

```

What we see is that the HMR model globally agrees with the `exp_zhao18` model in Fluxible.
However, when `selectfluxes` selects the robust linear fit, there is a mismatch.
Which makes sense, since it was demonstrated multiple times that exponential fits tends to overestimate fluxes, while linear fits tends to underestimate fluxes.
Let's see what happens if we are consistant.

```{r}
#| label: gasfluxes-comparison-hmr
#| echo: TRUE
#| fig-width: 8
#| fig-height: 9
#| fig-cap: ""

library(ggplot2)

gasfluxes_HMR <- gasfluxes_fits$result |>
    select(ID, "HMR.f0")

all_fluxes_HMR <- left_join(gasfluxes_HMR, fluxible_fluxes_hm, by = c("ID" = "f_fluxid"))

all_fluxes_HMR |>
    ggplot(aes(HMR.f0, fluxible_flux)) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0) +
    geom_smooth(
        method = "lm"
        ) +
    ggtitle("Fluxible exp_hm vs gasfluxes HMR")
    # geom_errorbar(aes(xmin = gasfluxes_flux - flux.se, xmax = gasfluxes_flux + flux.se))

```

Let's compare with the kappamax method in fluxible:
```{r}
#| label: gasfluxes-comparison-kappamax
#| echo: TRUE
#| fig-width: 8
#| fig-height: 9
#| fig-cap: ""
library(tidyverse)
library(fluxible)
library(ggplot2)


conc_liahovden <- flux_match(
  co2_liahovden,
  record_liahovden,
  datetime,
  start,
  conc,
  startcrop = 0,
  measurement_length = 180,
  ratio_threshold = 0.5,
  time_diff = 0
)

slopes_exp_liahovden <- flux_fitting(
  conc_liahovden,
  conc,
  datetime,
  end_cut = 10,
  fit_type = "exp_hm"
)

slopes_exp_liahovden_flag <- flux_quality(
  slopes_exp_liahovden,
  conc,
  f_rsquared = f_rsquared_lm,
  f_pvalue = f_pvalue_lm,
  instr_error = 5,
  kappamax = TRUE,
  error = 20
)

flux_plot(
    slopes_exp_liahovden_flag,
    conc,
    datetime,
    output = "pdfpages",
    f_plotname = "hm_kappamax3",
    f_ylim_lower = 300,
    )

fluxible_fluxes_kappamax <- flux_calc(
    slopes_exp_liahovden_flag,
    f_slope_corr, # we use the slopes provided by flux_quality
    datetime,
    temp_air,
    conc_unit = "ppm",
    flux_unit = "mmol",
    chamber_volume = 24.5,
    tube_volume = 0.075,
    atm_pressure = 1,
    plot_area = 0.0625,
    cols_keep = "type"
  ) |>
 select(f_fluxid, f_flux, type, f_model) |>
 rename(
    fluxible_flux = "f_flux"
 )


all_fluxes <- left_join(gasfluxes_fluxes, fluxible_fluxes_kappamax, by = c("ID" = "f_fluxid"))

all_fluxes |>
    ggplot(aes(gasfluxes_flux, fluxible_flux,
     color = f_model,
      label = ID)) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0) +
    geom_smooth(
        method = "lm"
        ) +
    geom_errorbar(aes(xmin = gasfluxes_flux - flux.se, xmax = gasfluxes_flux + flux.se)) +
    geom_text(hjust = 0, vjust = -1) +
    facet_grid(~method) +
    labs(
      title = "fluxible exp_hm kappamax vs gasfluxes kappamax",
      colour = "fluxible method"
    )

# investigating that flux with a massive difference

# slice_max(all_fluxes, fluxible_flux)

# flags_lia |>
# filter(f_fluxid == 46) |>
# flux_plot(conc, datetime)

```
Need to compare lm model
```{r}
#| label: data-process-stupeflux-lm
#| echo: TRUE
library(tidyverse)
library(fluxible)
library(ggplot2)

# we can use the stupeflux function, which is wrapper for fluxible
# we use the arguments that we already know from having processed
# those data in the detailed example.

fluxible_fluxes_lm <- stupeflux(
    raw_conc = co2_liahovden,
    field_record = record_liahovden,
    datetime_col = datetime,
    start_col = start,
    conc_col = conc,
    startcrop = 0,
    measurement_length = 180,
    fit_type = "linear",
    end_cut = 10,
    force_discard = c(51, 101),
    force_zero = 46,
    temp_air_col = temp_air,
    conc_unit = "ppm",
    flux_unit = "mmol",
    chamber_volume = 24.5,
    tube_volume = 0.075,
    atm_pressure = 1,
    plot_area = 0.0625,
    cols_keep = "type"
 ) |>
 select(f_fluxid, f_flux, type) |>
 rename(
    fluxible_flux = "f_flux"
 )

 gasfluxes_rlm <- gasfluxes_fits$result |>
    select(ID, "robust.linear.f0")

all_fluxes_HMR <- left_join(gasfluxes_rlm, fluxible_fluxes_lm, by = c("ID" = "f_fluxid"))

all_fluxes_HMR |>
    ggplot(aes(robust.linear.f0, fluxible_flux)) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0) +
    geom_smooth(
        method = "lm"
        ) +
    ggtitle("Fluxible linear vs gasfluxes robust linear")
```

So since both the hmr and linear fit agree, the problem is in finding kappamax.
```{r}
#| label: comparison-kappamax
#| echo: TRUE
#| fig-width: 8
#| fig-height: 9
#| fig-cap: ""
library(tidyverse)
library(fluxible)
library(ggplot2)


conc_liahovden <- flux_match(
  co2_liahovden,
  record_liahovden,
  datetime,
  start,
  conc,
  startcrop = 0,
  measurement_length = 180,
  ratio_threshold = 0.5,
  time_diff = 0
)

slopes_exp_liahovden <- flux_fitting(
  conc_liahovden,
  conc,
  datetime,
  end_cut = 10,
  fit_type = "exp_hm"
)

slopes_exp_liahovden_flag <- flux_quality(
  slopes_exp_liahovden,
  conc,
  f_rsquared = f_rsquared_lm,
  f_pvalue = f_pvalue_lm,
  instr_error = 5,
  kappamax = TRUE,
  error = 20
)

fluxible_kappamax <- slopes_exp_liahovden_flag |>
    select(f_fluxid, f_kappamax, f_model) |>
    distinct()

gasfluxes_kappamax <- selectfluxes(
                        gasfluxes_fits$result, # that is not specified in the vignette
                        select = "kappa.max",
                        f.detect = 0.031, # I keep the suggested value for now
                        t.meas = 0.05 # I think this is the length of the measurement in hours
                    ) |>
                    select(ID, kappa.max, method) |>
                    rename(g_kappamax = "kappa.max") |>
                    mutate(g_kappamax_abs = abs(g_kappamax))

all_kappamax <- left_join(gasfluxes_kappamax, fluxible_kappamax, by = c("ID" = "f_fluxid"))

all_kappamax |>
    ggplot(aes(g_kappamax, f_kappamax, label = ID, color = f_model)) +
    geom_point() +
    # geom_abline(slope = 1, intercept = 0) +
    geom_smooth(
        method = "lm"
        ) +
    geom_text(hjust = 0, vjust = -1) +
facet_grid(~method) +
    labs(
      title = "Fluxible kappamax vs gasfluxes kappamax",
      colour = "fluxible method"
    )
```
Now we see that the fluxes for which gasfluxes chose a linear fit while fluxible chose the exponential are those with a negative kappamax in gasfluxes.
Which makes sense, since kappa is force positive, so it will always be greater than a negative kappamax.
In the documentation of the `selectfluxes` function: "Note that the algorithm was developed for predominantly positive fluxes (such as N2O fluxes). If data with considerable gas uptake is analyzed, the algorithm needs to be modified, which currently means the user needs to implement it themselves."

What happens now if we take the absolute value of kappamax (like in fluxible)?
```{r}
#| label: gasfluxes-comparison-kappamax-absolute
#| echo: TRUE
#| fig-width: 8
#| fig-height: 9
#| fig-cap: ""
library(tidyverse)
library(fluxible)
library(ggplot2)


conc_liahovden <- flux_match(
  co2_liahovden,
  record_liahovden,
  datetime,
  start,
  conc,
  startcrop = 0,
  measurement_length = 180,
  ratio_threshold = 0.5,
  time_diff = 0
)

slopes_exp_liahovden <- flux_fitting(
  conc_liahovden,
  conc,
  datetime,
  end_cut = 10,
  fit_type = "exp_hm"
)

slopes_exp_liahovden_flag <- flux_quality(
  slopes_exp_liahovden,
  conc,
  f_rsquared = f_rsquared_lm,
  f_pvalue = f_pvalue_lm,
  instr_error = 5,
  kappamax = TRUE,
  error = 20
)

# flux_plot(
#     slopes_exp_liahovden_flag,
#     conc,
#     datetime,
#     output = "pdfpages",
#     f_plotname = "hm_kappamax3",
#     f_ylim_lower = 300,
#     )

fluxible_fluxes_kappamax <- flux_calc(
    slopes_exp_liahovden_flag,
    f_slope_corr, # we use the slopes provided by flux_quality
    datetime,
    temp_air,
    conc_unit = "ppm",
    flux_unit = "mmol",
    chamber_volume = 24.5,
    tube_volume = 0.075,
    atm_pressure = 1,
    plot_area = 0.0625,
    cols_keep = "type"
  ) |>
 select(f_fluxid, f_flux, type, f_model) |>
 rename(
    fluxible_flux = "f_flux"
 )

gasfluxes_fluxes_abs <- gasfluxes_fits$result |>
    mutate(kappa.max = abs(kappa.max)) |>
    selectfluxes(select = "kappa.max",
                        f.detect = 0.031, # I keep the suggested value for now
                        t.meas = 0.05 # I think this is the length of the measurement in hours
                    ) |>
                    select(ID, flux, method, flux.se) |>
                    rename(
                        gasfluxes_flux = "flux"
                    )


all_fluxes <- left_join(gasfluxes_fluxes_abs, fluxible_fluxes_kappamax, by = c("ID" = "f_fluxid"))

all_fluxes |>
    ggplot(aes(gasfluxes_flux, fluxible_flux,
     color = f_model,
      label = ID)) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0) +
    geom_smooth(
        method = "lm"
        ) +
    geom_errorbar(aes(xmin = gasfluxes_flux - flux.se, xmax = gasfluxes_flux + flux.se)) +
    geom_text(hjust = 0, vjust = -1) +
    facet_grid(~method) +
    labs(
      title = "fluxible exp_hm kappamax vs gasfluxes abs(kappamax)",
      colour = "fluxible method"
    )

```
That did not show anything.
Let's compare b and kappa
```{r}
#| label: comparison-kappa-b
#| echo: TRUE
#| fig-width: 8
#| fig-height: 9
#| fig-cap: ""
library(tidyverse)
library(fluxible)
library(ggplot2)


conc_liahovden <- flux_match(
  co2_liahovden,
  record_liahovden,
  datetime,
  start,
  conc,
  startcrop = 0,
  measurement_length = 180,
  ratio_threshold = 0.5,
  time_diff = 0
)

slopes_exp_liahovden <- flux_fitting(
  conc_liahovden,
  conc,
  datetime,
  end_cut = 10,
  fit_type = "exp_hm"
)

slopes_exp_liahovden_flag <- flux_quality(
  slopes_exp_liahovden,
  conc,
  f_rsquared = f_rsquared_lm,
  f_pvalue = f_pvalue_lm,
  instr_error = 5,
  kappamax = TRUE,
  error = 20
)

fluxible_b <- slopes_exp_liahovden_flag |>
    select(f_fluxid, f_b, f_model) |>
    distinct()

gasfluxes_kappa <- selectfluxes(
                        gasfluxes_fits$result, # that is not specified in the vignette
                        select = "kappa.max",
                        f.detect = 0.031, # I keep the suggested value for now
                        t.meas = 0.05 # I think this is the length of the measurement in hours
                    ) |>
                    select(ID, HMR.kappa, method)

all_kappa <- left_join(gasfluxes_kappa, fluxible_b, by = c("ID" = "f_fluxid"))

all_kappa |>
    ggplot(aes(HMR.kappa, f_b, label = ID
    # , color = f_model
    )) +
    geom_point() +
    # geom_abline(slope = 1, intercept = 0) +
    geom_smooth(
        method = "lm"
        ) +
    geom_text(hjust = 0, vjust = -1) +
# facet_grid(~method) +
    labs(
      title = "Fluxible b (kappa) vs gasfluxes kappa"
    #   colour = "fluxible method"
    )
```

Ok, we need to redo all of this with only the fluxes that disagreed.
```{r}
#| label: comparison-different-model-only
#| echo: TRUE
#| fig-width: 8
#| fig-height: 9
#| fig-cap: ""
library(tidyverse)
library(fluxible)
library(ggplot2)

conc_liahovden <- flux_match(
  co2_liahovden,
  record_liahovden,
  datetime,
  start,
  conc,
  startcrop = 0,
  measurement_length = 180,
  ratio_threshold = 0.5,
  time_diff = 0
)

slopes_exp_liahovden <- flux_fitting(
  conc_liahovden,
  conc,
  datetime,
  end_cut = 10,
  fit_type = "exp_hm"
)

slopes_exp_liahovden_flag <- flux_quality(
  slopes_exp_liahovden,
  conc,
  f_rsquared = f_rsquared_lm,
  f_pvalue = f_pvalue_lm,
  instr_error = 5,
  kappamax = TRUE,
  error = 20
)

# flux_plot(
#     slopes_exp_liahovden_flag,
#     conc,
#     datetime,
#     output = "pdfpages",
#     f_plotname = "hm_kappamax3",
#     f_ylim_lower = 300,
#     )

fluxible_fluxes_kappamax <- flux_calc(
    slopes_exp_liahovden_flag,
    f_slope_corr, # we use the slopes provided by flux_quality
    datetime,
    temp_air,
    conc_unit = "ppm",
    flux_unit = "mmol",
    chamber_volume = 24.5,
    tube_volume = 0.075,
    atm_pressure = 1,
    plot_area = 0.0625,
    cols_keep = c("type", "f_b", "f_kappamax", "f_slope_lm")
  ) |>
 select(f_fluxid, f_flux, f_b, f_kappamax, type, f_model, f_slope_lm) |>
 rename(
    fluxible_flux = "f_flux"
 ) |>
 filter(fluxible_flux >= 0)

gasfluxes_fluxes <- selectfluxes(
                        gasfluxes_fits$result, # that is not specified in the vignette
                        select = "kappa.max",
                        f.detect = 0.031, # I keep the suggested value for now
                        t.meas = 0.05 # I think this is the length of the measurement in hours
                    ) |>
                    select(ID, flux, HMR.kappa, kappa.max, method, flux.se, robust.linear.f0) |>
                    rename(
                        gasfluxes_flux = "flux",
                        gasfluxes_kappa = "HMR.kappa",
                        gasfluxes_kappamax = "kappa.max",
                        gasfluxes_lm = "robust.linear.f0"
                    ) |>
                    filter(gasfluxes_flux >= 0)

all_fluxes <- left_join(gasfluxes_fluxes, fluxible_fluxes_kappamax, by = c("ID" = "f_fluxid")) |>
    mutate(
    f_model = str_replace_all(f_model, "exp_hm", "exp"),
    method = str_replace_all(method, c("HMR" = "exp", "robust linear" = "linear"))
    )

prob_fluxes <- all_fluxes |>
    filter(f_model != method)

# same plot as before
prob_fluxes |>
    ggplot(aes(gasfluxes_flux, fluxible_flux,
     color = f_model,
      label = ID)) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0) +
    geom_smooth(
        method = "lm"
        ) +
    geom_errorbar(aes(xmin = gasfluxes_flux - flux.se, xmax = gasfluxes_flux + flux.se)) +
    geom_text(hjust = 0, vjust = -1) +
    # facet_grid(~method) +
    labs(
      title = "fluxible exp_hm kappamax vs gasfluxes kappamax",
      subtitle = "different models only",
      colour = "fluxible method"
    )

# now let's look at kappa
prob_fluxes |>
    ggplot(aes(gasfluxes_kappa, f_b,
     color = f_model,
      label = ID)) +
    geom_point() +
    # geom_abline(slope = 1, intercept = 0) +
    geom_smooth(
        method = "lm"
        ) +
    # geom_errorbar(aes(xmin = gasfluxes_flux - flux.se, xmax = gasfluxes_flux + flux.se)) +
    geom_text(hjust = 0, vjust = -1) +
    # facet_grid(~method) +
    labs(
      title = "fluxible b (kappa) vs gasfluxes kappa",
      subtitle = "different models only",
      colour = "fluxible method"
    )
# and kappamax
prob_fluxes |>
    ggplot(aes(gasfluxes_kappamax, f_kappamax,
     color = f_model,
      label = ID)) +
    geom_point() +
    # geom_abline(slope = 1, intercept = 0) +
    geom_smooth(
        method = "lm"
        ) +
    # geom_errorbar(aes(xmin = gasfluxes_flux - flux.se, xmax = gasfluxes_flux + flux.se)) +
    geom_text(hjust = 0, vjust = -1) +
    # facet_grid(~method) +
    labs(
      title = "fluxible kappamax vs gasfluxes kappamax",
      subtitle = "different models only",
      colour = "fluxible method"
    )

# and lm slope
prob_fluxes |>
    ggplot(aes(gasfluxes_lm, f_slope_lm,
     color = f_model,
      label = ID)) +
    geom_point() +
    # geom_abline(slope = 1, intercept = 0) +
    geom_smooth(
        method = "lm"
        ) +
    # geom_errorbar(aes(xmin = gasfluxes_flux - flux.se, xmax = gasfluxes_flux + flux.se)) +
    geom_text(hjust = 0, vjust = -1) +
    # facet_grid(~method) +
    labs(
      title = "fluxible lm vs gasfluxes robust linear",
      subtitle = "different models only",
      colour = "fluxible method"
    )

```
